---
title: "Presentation ready plots"
author: "MACS 40700 <br /> University of Chicago"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      highlightStyle: magula
      highlightLines: true
      highlightLanguage: r
      ratio: 16:9
      countIncrementalSlides: false
      
---

```{r setup, include = FALSE, cache = FALSE}
# generate CSS file
library(xaringanthemer)
style_xaringan(
  text_color = "#464A53",
  header_color = "#586891",
  background_color = "#F1F3F2",
  title_slide_background_color = "#586891",
  title_slide_text_color = "#F1F3F2",
  table_row_even_background_color = "#B3A6A3",
  link_color= "#586891",
  text_bold_color= "#586891",
  #code_highlight_color= rgba(255,255,0,0.5),
  inverse_text_color= "#464A53",
  inverse_background_color= "#B3A6A3",
  inverse_header_color= "#464A53",
  inverse_link_color= "#586891",
  header_background_color= "#586891",
  header_background_text_color= "#F1F3F2",
  code_highlight_color = "#9ab7c4",
  text_font_family = "Nunito",
  text_font_url = "https://fonts.google.com/specimen/Nunito?stroke=Sans+Serif",
  header_font_google = google_font("Afacad"),
  base_font_size= "28px",
  text_font_size= "1rem",
  code_font_size= "20px",
  footnote_font_size = "0.75em",
  code_inline_font_size= "1em",
  header_h1_font_size= "2rem",
  header_h2_font_size= "1.75rem",
  header_h3_font_size= "1.5rem")

# source in the default knitr options
source(here::here("R", "slide-opts.R"))

# enable panelsets and default theme
xaringanExtra::use_panelset()
# ggplot2::theme_set(ggplot2::theme_minimal(base_size = 16))

# load basic packages
library(knitr)
library(here)
library(countdown)
library(patchwork)
```

class: middle, inverse

# Agenda:

* Wrap up annotations
* Telling a story
* `xaringan`
* Plots on slides


---

## Announcements 

---

## Setup

```{r message = FALSE, warning = FALSE}
# load packages
library(tidyverse)
library(colorspace)
library(ggtext)
library(scales)
library(ggrepel)

# set default theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 16))

# set default figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 8, fig.asp = 0.618, fig.retina = 3,
  dpi = 300, out.width = "60%"
)

# dplyr print min and max
options(dplyr.print_max = 6, dplyr.print_min = 6)
```

---

class: inverse, middle

# Telling a story

---

## Multiple ways of telling a story

- Sequential plots: Motivation, then resolution

- A single plot: Resolution, and hidden in it motivation

<br>

.note[
**Project note:** you're asked to create three plots. One possible approach: Start with a plot showing the raw data, and show derived quantities (e.g. percent increases, averages, coefficients of fitted models) in the subsequent plot, and some combination in the third. 
]

---

## Simplicity vs. complexity

> When you’re trying to show too much data at once you may end up not showing anything.

- Never assume your audience can rapidly process complex visual displays

- Don't add variables to your plot that are only tangential to your story

- Don't jump straight to a highly complex figure; first show an easily digestible subset (e.g., show one facet first)

- Aim for memorable, but clear


---

## Consistency vs. repetitiveness

> Be consistent but don’t be repetitive.

- Use consistent features throughout plots (e.g., same color represents same level on all plots)

- Aim to use a different type of visualization for each distinct analysis

---

class: inverse, middle

# `xaringan`

---

## `xaringan`

- R Markdown extension to HTML5 presentations

- [Detailed guide to `xaringan`](https://arm.rbind.io/slides/xaringan.html)

- [`xaringan` in R Markdown reference book](https://bookdown.org/yihui/rmarkdown/xaringan.html)

- [Example slides](https://slides.yihui.org/xaringan/#1) 

<!-- #https://dataviz-sp22.github.io/proj1/presentation/presentation.html#1) -->

---

class: inverse, middle

# Plots on slides

---

## Fitting your figures on a slide

- Use a setup chunk for consistent `dpi`, `fig.asp`, `out.width` etc.

- But, remember that you might want different `fig.asp` and `out.width` for faceted plots

- For a "zoomed out" look, increase `fig.width`

.pull-left[
```{r out.width = "100%", echo = FALSE}
mtcars %>%
  rownames_to_column() %>%
  ggplot(aes(x = disp, y = mpg)) +
  geom_text_repel(aes(label = rowname)) +
  coord_cartesian(clip = "off") +
  labs(title = "fig.width = 8")
```
]
.pull-right[
```{r out.width = "100%", fig.width = 10, echo = FALSE}
mtcars %>%
  rownames_to_column() %>%
  ggplot(aes(x = disp, y = mpg)) +
  geom_text_repel(aes(label = rowname)) +
  coord_cartesian(clip = "off") +
  labs(title = "fig.width = 10")
```
]

---

## Multiple figures on a slide

- Use `.pull-left` and `.pull-right` environments  

- Use `out.width= "50%"` in a chunk that creates two plots to have each span half the slide and be placed next to each other (or `out.width = "33%"` for three plots, etc.)

- Use the [**patchwork**](https://patchwork.data-imaginist.com/) package (details on this later in the quarter, but feel free to look into it yourself!)

- Possibly, use pivoting to reshape your data and then use facets

---

<!-- ## Like something you saw in my slides? -->

<!-- <br> -->

<!-- Look into the source code at https://github.com/uc-dataviz/slides. -->

# Try it out

* Create a four-slide grouping of your final project so far
* Aim to have three-ish bullet points per slide
