---
title: "Plotly Party"
author: "MACS 40700 <br /> University of Chicago"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      highlightStyle: magula
      highlightLines: true
      highlightLanguage: r
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include = FALSE, cache = FALSE}
# generate CSS file
library(xaringanthemer)
style_xaringan(
  text_color = "#464A53",
  header_color = "#586891",
  background_color = "#F1F3F2",
  title_slide_background_color = "#586891",
  title_slide_text_color = "#F1F3F2",
  table_row_even_background_color = "#B3A6A3",
  link_color= "#586891",
  text_bold_color= "#586891",
  #code_highlight_color= rgba(255,255,0,0.5),
  inverse_text_color= "#464A53",
  inverse_background_color= "#B3A6A3",
  inverse_header_color= "#464A53",
  inverse_link_color= "#586891",
  header_background_color= "#586891",
  header_background_text_color= "#F1F3F2",
  code_highlight_color = "#9ab7c4",
  text_font_family = "Nunito",
  text_font_url = "https://fonts.google.com/specimen/Nunito?stroke=Sans+Serif",
  header_font_google = google_font("Afacad"),
  base_font_size= "28px",
  text_font_size= "1rem",
  code_font_size= "20px",
  footnote_font_size = "0.75em",
  code_inline_font_size= "1em",
  header_h1_font_size= "2rem",
  header_h2_font_size= "1.75rem",
  header_h3_font_size= "1.5rem")

source(here::here("R", "slide-opts.R"))

xaringanExtra::use_panelset()
ggplot2::theme_set(ggplot2::theme_gray(base_size = 16))
```

```{r pkgs, include = FALSE, cache = FALSE}
library(knitr)
library(here)
library(tidyverse)
library(plotly)
```

class: middle, inverse

# Data visualization: Agenda for today


---

# Plotly

plotly is an R package for creating interactive web-based graphs via the open source JavaScript graphing library plotly.js.


As of version 2.0 (November 17, 2015), graphs created with the plotly R package are rendered locally through the htmlwidgets framework.


Plotly graphs are great as they have a nice interactive/mouseover component. 



---

# Plotly: getting started

```{r, eval = FALSE}
install.packages("plotly")
```

```{r}
library(plotly)

fig <- plot_ly(midwest, x = ~percollege, color = ~state, type = "box", height = 250) 
fig
```

---

# Plotly: options

* Bar plots
* Scatter plots
* Line plots
* Dot plots
* Barbel plots
* Boxplots



---
# Bar

```{r}
fig <- plot_ly(
  x = c("giraffes", "orangutans", "monkeys"),
  y = c(20, 14, 23),
  name = "SF Zoo",
  type = "bar", height = 150
)

fig
```

---
# Scatter plots
```{r}
plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, height = 400)
```


---
## Line plots

```{r}
trace_0 <- rnorm(100, mean = 5)
trace_1 <- rnorm(100, mean = 0)
trace_2 <- rnorm(100, mean = -5)
x <- c(1:100)

data <- data.frame(x, trace_0, trace_1, trace_2)

fig <- plot_ly(data, x = ~x, height = 200)
fig <- fig %>% add_trace(y = ~trace_0, name = 'trace 0',mode = 'lines')
fig <- fig %>% add_trace(y = ~trace_1, name = 'trace 1', mode = 'lines+markers')
fig <- fig %>% add_trace(y = ~trace_2, name = 'trace 2', mode = 'markers')

fig
```

---
## Dot plots
```{r}
p <-    
 ggplot(mtcars, aes(x = mpg, fill = factor(cyl))) +
  geom_dotplot(stackgroups = TRUE, binwidth = 1, binpositions = "all")

ggplotly(p, height = 325) 


```

---
##  Barbel/Dumbell plots

.panelset[
.panel[.panel-name[Dumbell plots]
```{r dumbell, fig.show = "hide"}
gapminder <- read_csv("https://raw.githubusercontent.com/datavizpyr/data/master/gapminder-FiveYearData.csv")
df <- gapminder %>%
  filter(year %in% c(1952,2007)) %>%
  filter(continent=="Asia") 

df <- df %>%
  mutate(paired = rep(1:(n()/2),each=2),
         year=factor(year))

p <- df %>% 
  group_by(paired) %>%
  ggplot(aes(x= lifeExp, y= reorder(country,lifeExp))) +
  geom_line(aes(group = paired),color="grey")+
    geom_point(aes(color=year), size=2) +
  labs(y="country") + theme(text = element_text(size = 10))

ggplotly(p, height = 450)
```
]
.panel[.panel-name[Plot]
```{r ref.label="dumbell", echo = FALSE, out.width="70%"}
```
]
]

---
## Colorscales 
Just like in ggplot, you can color by variables. 

```{r}
plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, color = ~Species, height = 300)

```

---
## Colorscales 

```{r}
library(palmerpenguins)
plot_ly(data = penguins, x = ~bill_length_mm , y = ~body_mass_g, color = ~sex,
        height = 300)

```

---
## ColorBrewer Palette Names
```{r}
plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, color = ~Species, 
        colors = "Set1", height = 300)

```

---
## Custom Color Scales
```{r}
pal <- c("red", "blue", "orange")

plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, color = ~Species, 
        colors = pal, height = 300)

```

---
## Mapping Data to Symbols
```{r}
plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, type = 'scatter',
  mode = 'markers', symbol = ~Species, symbols = c('circle','x','o'),
  color = I('black'), marker = list(size = 10), height = 300)
```

---
## Data labels on hover
```{r}
d <- diamonds[sample(nrow(diamonds), 1000), ]

plot_ly(
  d, x = ~carat, y = ~price,
  # Hover text:
  text = ~paste("Price: ", price, '$<br>Cut:', cut),
  color = ~carat, size = ~carat, height = 300)
 
```


---
# Boxplots

```{r}
set.seed(1234)
dat <- data.frame(cond = factor(rep(c("A","B"), each=200)), rating = c(rnorm(200),rnorm(200, mean=.8)))

p <- ggplot(dat, aes(x=cond, y=rating, fill=cond)) + geom_boxplot() 

ggplotly(p, height = 350)  
```

---
## Data labels on hover
```{r}
d <- diamonds[sample(nrow(diamonds), 1000), ]

plot_ly(
  dat, x = ~cond, y = ~rating,
  # Hover text:
  text = ~paste("Price: ", cond),
  color = ~cond, type = "box", height = 200)
```

---

## Additional plot types

* Just about any other plot type can be done -- try it on your own
  * Animated plots
  * Area plots
  * Bubble plots
  * Contour plots
  * Heatmaps
  * Spatial maps
  * Etc. 

---

# Recap

* Plotly is another way to plot things as it offers interactivity. It can be used in R or Python and the plots can be quite nice if you're making slides or anything that's going to be online in some way. 

* Sometimes you can incorporate ggplot elements and then render in ggplot

* Just about any plot you make in ggplot you can make in plotly
  * Boxplots
  * Bar plots
  * Scatter plots
  * Line plots
  * ... etc. ...
