---
date: "2022-09-26T00:00:00-05:00"
draft: false
weight: 40
title: "Accessing RStudio Workbench"
toc: true
type: book
aliases: ["/setup_server.html", "/setup/r-server/"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```
Let's begin with a few definitions:

## R

[R](https://www.r-project.org/) is an open-source programming language based on the [S](https://en.wikipedia.org/wiki/S_(programming_language)) from the 1970s. It is very popular in the physical and social sciences due to it's cost (free) and versatility. Thousands of expansion libraries have been published which extend the tasks R can perform, and users can write their own custom functions and/or libraries to perform specific operations.

## RStudio

The base R distribution is not the best for developing and writing programs. Instead, we want an integrated development environment (IDE) which will allow us to write and execute code, debug programs, and automate certain tasks. In this course we will use [RStudio](https://www.rstudio.com/products/RStudio/), perhaps the most popular IDE available for R. Like R, it is open-source, expandable, and provides many useful tools and enhancements over the base R environment.

We will be using RStudio for this course so that you can practice a workflow you will continue. 



## Acknowledgments

```{r child = here::here("R", "_ack_stat545.Rmd")}
```
```{r child = here::here("R", "_ack_ben.Rmd")}
```